---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
Sys.setlocale("LC_ALL","English")
```

## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?
```{r}
library(dplyr)

stepsPerDay <- data %>%
  group_by(date) %>%
  summarise(total = sum(steps))

hist(stepsPerDay$total, main="Steps taken per day", xlab="Total steps taken per day", breaks=15)
```
Mean of the total number of steps taken per day: 
```{r}
mean(stepsPerDay$total, na.rm=TRUE)
```


Median of the total number of steps taken per day: 
```{r}
median(stepsPerDay$total, na.rm=TRUE)
```


## What is the average daily activity pattern?
```{r}
meanSteps <- aggregate(steps ~ interval, data, mean)

plot(meanSteps$interval, meanSteps$steps,  type = "l",  main="Mean steps", xlab="Interval", ylab="Average steps taken")
```
5-minute interval containing the maximum number of steps:

```{r}
meanSteps[which(meanSteps$steps == max(meanSteps$steps)), ]
```

## Imputing missing values

Number of missing values:

```{r}
sum(is.na(data))
```

Impute missing values with the mean of the corresponding interval:

```{r}
dataImputed <- data

for (i in 1:nrow(dataImputed)) {
  if (is.na(dataImputed[i, 1])) {
    dataImputed[i, 1] <- meanSteps[meanSteps$interval %in% dataImputed[i, 3], 2]
  }
}
```

Check number of NA values:

```{r}
sum(is.na(dataImputed))
```

```{r}

stepsPerDayImputed <- dataImputed %>%
  group_by(date) %>%
  summarise(total = sum(steps))

hist(stepsPerDayImputed$total, main="Steps taken per day imputed", xlab="Total steps taken per day", breaks=15)
```

Mean of the total number of steps taken per day: 
```{r}
mean(stepsPerDayImputed$total, na.rm=TRUE)
```


Median of the total number of steps taken per day: 
```{r}
median(stepsPerDayImputed$total, na.rm=TRUE)
```


There is a slight difference in the median.


## Are there differences in activity patterns between weekdays and weekends?


```{r}
# Create gender vector
dataImputed$day <- weekdays(as.Date(dataImputed$date))

for(i in 1:nrow(dataImputed)) {
  if (dataImputed[i, 4] %in% c("Saturday", "Sunday")) {
    dataImputed[i, 4] <- "weekend"
  } else {
    dataImputed[i, 4] <- "weekday"
  }
}

dataImputed$day <- factor(dataImputed$day)

head(dataImputed)

```

```{r}
weekdaySteps <- dataImputed[dataImputed$day == 'weekday', ]
weekendSteps <- dataImputed[dataImputed$day == 'weekend', ]

weekdayStepsMean <- aggregate(steps ~ interval, weekdaySteps, mean)
weekendStepsMean <- aggregate(steps ~ interval, weekendSteps, mean)

head(weekdayStepsMean)

head(weekendStepsMean)

```

```{r}

par(mfcol=c(1, 2))


plot(weekdayStepsMean$interval, weekdayStepsMean$steps, type="l", xlab="interval", ylab="steps", main="weekdays")

plot(weekendStepsMean$interval, weekendStepsMean$steps, type="l", xlab="interval", ylab="steps", main="weekends")



```
